---
title: "An Introduction to Visibly"
author: "Michael Clark"
date: <span style="font-style:normal;font-family:'Open Sans'">`r Sys.Date()`</span>
output:
  html_vignette:
    toc: true
    toc_depth: 3
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, collapse = TRUE,
                      comment=NA, R.options=list(width=220),   # code 
                      dev.args=list(bg = 'transparent'), dev='png',                                 # viz
                      fig.align='center', out.width='75%', fig.asp=.75,                 
                      cache.rebuild=F, cache=F)                                                     # cache
```




## Palettes & Themes

One can get started by creating a palette. This uses the <span class="pack">colortools</span> package to provide evenly spaced colors, given a starting color.

```{r example, echo=-(1:3)}
# bc r cmd build randomly decided margins were too large when they were okay
# before
par(mar=c(.5,.5,.5,.5))
library(visibly)
create_palette('papayawhip', plot = TRUE)
```

Here is a palette based on the blue in the R logo.

```{r example2}
palettes$Rblue
```


The palette creator can create some decent categorical distinctions without too much fuss. The following also demonstrates one of the themes, which has no grid/gray, and de-bolds the black font while leaving text clear; even the fainter version will pass web standards for contrast against a white background. As shown, you can still fiddle with the theme beyond that.

```{r use1}
pal = create_palette('#ff5500', 
                     name = 'orange_you_glad_you_have_this_color')
library(ggplot2)

ggplot(mtcars, aes(x=wt, y=mpg)) +
  geom_point(aes(color=factor(cyl)), size=10, alpha=.5) +
  scale_color_manual(values = pal$triadic) +
  theme_trueMinimal()
```

```{r use2}
library(dplyr)
mtcars %>% 
  mutate(cyl = factor(cyl)) %>% 
  tidyext::num_by(wt, cyl) %>% 
  ggplot(aes(x=cyl, y=Mean)) +
  geom_col(aes(fill=cyl), width=.5, alpha=.85) +
  scale_fill_manual(values = palettes$Rblue$triadic) +
  theme_trueMinimal() +
  theme(legend.key.size = unit(.015, 'npc'),
        axis.title.y = element_text(size=20, hjust=-.05))
```


Another way to create palettes is with a function that works with [colorgorical](http://vrl.cs.brown.edu/color).  It connects to that website to create colors based on things like whether they perceptually go together aesthetically, whether they are more or less perceptually distinct, or even if their name is unique.

```{r colorgorical, eval=2:4, echo=1}
colorgorical(n=6, pairPreference = 1, startPalette = list(c(10, -60, 45)), output = 'hex')
x = c("#002B00", "#95C857", "#334D37", "#4EF185", "#378811", "#7FE7D3")
print(x)
qplot(x=x, y=1:6, color=I(x), size=I(10))
```

```{r colorgorical2, eval=2:4, echo=1}
colorgorical(n=10, perceptualDifference = .5, startPalette = list(c(10, -60, 45)), output = 'hex')
x = c("#002B00", "#D57381", "#77CE3F", "#DB0EAC", "#2FF52B", 
      "#6C208E", "#B1BF81", "#4115F9", "#518512", "#B662CA")
print(x)
qplot(x=x, y=1:10, color=I(x), size=I(10))
```

To get a starting palette from an R or hexadecimal value, try <span class="pack">col2lab</span>.

```{r col2lab}
col2lab('dodgerblue')
```

You can check your to see if your chosen color combinations would pass web standards for contrast sensitivity.  For example, if you use blue for your hyperlinks and consider it against a white background for a webpage it passes.

```{r color_contrast_checker}
color_contrast_checker(foreground = 'blue')
```

However, if you consider the background as the surrounding text, perhaps not.

```{r color_contrast_checker2}
color_contrast_checker(foreground = 'blue', background = 'black')
```

Default ggplot2 colors don't contrast well with the background. Ack!

```{r color_contrast_checker3}
iris %>% 
  ggplot(aes(Petal.Length, Petal.Width, color=Species)) + 
  geom_point()

scales::show_col(scales::hue_pal()(3))

color_contrast_checker(foreground = '#F8766D', background = 'gray92')
color_contrast_checker(foreground = '#00BA38', background = 'gray92')
color_contrast_checker(foreground = '#619CFF', background = 'gray92')
```



However, if you want palettes that are colorblind-safe, print-safe etc., especially for continuous scales, you should use packages like [viridis](https://github.com/sjmgarnier/viridis) and [scico](https://github.com/thomasp85/scico).



## Descriptive Visualization

People often want to (and should) examine correlation matrices as a prelude to modeling.  One problem is that it is difficult to make much sense from the matrix when there are more than a few columns in the original data, as there are too many pairwise values present.  However a little reshuffling of the rows/columns, and visually depicting the values by color can make patterns pop out quickly.

You have plenty of options for this sort of endeavor, such as <span class="pack">corrplot</span>, <span class="pack">psych</span>, and <span class="pack">heatmaply.</span>  However, they offer only static plots, or reorder based on a cluster analysis.  I originally had [a function](https://github.com/m-clark/heatR) that would provide some interactivity, but while utilizing a factor analytic approach. I've now deprecated that standalone package and brought the functionality here. The <span class="pack">corr_heat</span> function provides an easy way to visualize your correlations.

```{r corrheat1}
data('bfi', package = 'visibly')
cor_matrix = cor(bfi, use='pair')
corr_heat(cor_matrix)
```

The <span class="pack">psych</span> package is used to for the factor analysis, which is better suited to survey data and similar settings, where items are of the same type, and you can supply arguments to its <span class="func">fa</span> function. The plot comes from <span class="pack">plotly</span>, and hovering over the squares provides the value.

```{r corrheat2}
corr_heat(cor_matrix, n_factors = 2, psych_opts = list(fm='ml', rot='oblimin'))
```

The color scheme is based on the previously mentioned scico package.

```{r corrheat3}
corr_heat(cor_matrix, pal = 'broc')
```


We can see that while there might be the usual 'Big Five' personality structure, it may also really just be two underlying constructs- neuroticism vs. the rest, at least for this sample.

It occurred to me while updating this functionality that a 3-d approach might actually be useful here (it isn't the vast majority of the time).  When looking at the previous structure in 3-d, we probably would only think about two factors of items.

```{r corrheat3d}
p = corr_heat(cor_matrix, three_d = TRUE, diagonal = NA)
p
```


There are multiple orderings available to help bring out the structure.  This may be particularly useful if you are not dealing with measurement error type data.

```{r corrheat4}
p = corr_heat(cor(mtcars), three_d = TRUE, ordering = 'polar')
p
```

Who knew the <span class="objclass">mtcars</span> data set could actually be interesting!




## Plotting Model Results

I have some visualizations for plotting uncertainty intervals for fixed and random effects of mixed models.  These require the <span class="pack">scico</span> package, as well as <span class="pack">lme4</span> and <span class="pack">brms</span> which are used to do the mixed models in the first place.

### lm and glm

I don't really use them aside for demonstration, but in this case the <span class="objclass">lm</span> and <span class="objclass">glm</span> objects served as a baseline.

```{r lm0}
fit_lm = lm(mpg ~ ., mtcars)
plot_coefficients(fit_lm)
```

Change the palette, order and more.

```{r lm1}
fit_lm = lm(mpg ~ ., mtcars)
plot_coefficients(fit_lm, 
                  palette='oslo', 
                  order = 'decreasing', 
                  sd_multi = 1,
                  keep_intercept = TRUE,
                  ref_line = c(-1:1))
```

#### Data only

You may want to do your own visualization, or go beyond the default settings. Rather than providing a couple dozen arguments for you to tweak, just so you can still feel like you can't get it just quite right.  Here's the data, do with it what you will!

```{r data}
plot_coefficients(fit_lm, plot=FALSE)
```


### Mixed models

I do a lot of mixed models, which is my main reason for providing this functionality, so I've started to create some ways to plot the results of the separate fixed and random effects.  We'll use <span class="pack">lme4</span>.

```{r fe1}
library(lme4)
fit_mer = lmer(Reaction ~ Days + (Days|Subject), sleepstudy)
plot_coefficients(fit_mer)
```

And now the random effects. Fewer options can be found here, but this is intentional as you'd want them ordered and the fixed effect style would make things messy whenever there are many random effects, which is often the case. With multiple random effects, a list of ggplot objects is returned.

```{r re1}
plot_coefficients(fit_mer, ranef = TRUE, which_ranef = 'Subject')
```

Now for a different palette.  Again, these only apply to fixed effects visualizations, due to the number of categories typically seen with random effects and how the visualization is created.

```{r fe2}
fit_mer2 = lmer(count ~ zAge + zBase * Trt + 
                  (1 | patient), 
                data = brms::epilepsy)
plot_coefficients(fit_mer2, 
                  palette = 'berlin')
```

You can use <span class="pack">patchwork</span> to put them into one graph.

```{r re_patchwork}
plots = plot_coefficients(fit_mer, ranef = TRUE, which_ranef = 'Subject')

library(patchwork)

plots[[1]] + plots[[2]]
```




#### brms

I like <span class="pack">brms</span> quite a bit and use it a lot. It served as the motivation for these coefficient plots, as did <span class="pack">bayesplot</span>, and <span class="pack">tidyposterior</span>.

```{r fe_brms1, eval=FALSE}
library(brms)
fit_brms = brm(count ~ log_Age_c + log_Base4_c * Trt + 
                 (1 | patient) + (1 | obs), 
               data = epilepsy,
               family = poisson)
plot_coefficients(fit_brms)
```

```{r fe_brms1_vis, echo=FALSE}
load('../tests/testthat/brms_res.RData')
plot_coefficients(fit1)
```

And now the random effects.

```{r re_brms1}
plot_coefficients(fit1, ranef = TRUE, which_ranef = 'patient') +
  theme(axis.text.x = element_text(angle = -90))
```

### GAM

Some very rudimentary plotting of basic GAM effects provided by the <span class="pack">mgcv</span> are provided.  I really like the package, but am not crazy for the default plots, and some offered by some other packages are okay, but may not be polished enough or do the smooth plots I'd eventually like to do.

```{r gam}
library(mgcv)
d = gamSim()

gam_model = gam(y ~ x0 + s(x1) + s(x2, bs='gp') + s(x3, bs='ps'), data=d)

plot_gam(gam_model,
         main_var = x2)

plot_gam(gam_model, 
         main_var = vars(x1, x2, x3),
         ncol = 1, 
         line_color = palettes$Rblue$Rblue, 
         ribbon_color = palettes$Rblue$complementary[2])

plot_gam_check(gam_model)
```



#### 2d smooths

Some initial functionality is provided for continuous-by-continuous interactions and continuous-by-categorical interaction plots.

```{r gam2d, echo=-1}
set.seed(1234)
d = gamSim(2, scale = .1)$data
gam_model = gam(y ~ s(x, z, bs='gp'), data=d)

plot_gam_2d(gam_model,
            main_var = x,
            second_var = z,
            n_plot = 200)
```


In 3d.

```{r gam3d}
plot_gam_3d(gam_model,
            main_var = x,
            second_var = z,
            n_plot = 200)
```



A categorical moderator.

```{r gamby, echo=-1}
set.seed(1234)
d = gamSim(4)
gam_model = gam(y ~ s(x2, fac, bs='fs'), data=d)

plot_gam_by(gam_model,
            main_var = x2,
            by_var = fac)
```

